---
title: "RWorksheet4c"
author: "Jonash Lor Defensor"
date: "2023-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  Download and open the mpg file. Upload it to your OWN environment
1. a. Show your solutions on how to import a csv file into the environment.

```{r}
library(readr)

open_mpg <- read.csv("mpg.csv")
open_mpg
```

1b. Which variables from mpg dataset are categorical

```{r}
str(open_mpg)
```
ans. manufacturer, model, trans, drv, fl, class variables are categorical

1c. Which are continuous variables?
ans. X, displ, year, cyl, cty, hwy are continuous variables

2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.

```{r}

manufacturer_counts <- table(open_mpg$manufacturer)



most_models_manufacturer <- as.character(names(manufacturer_counts)[which.max(manufacturer_counts)])


model_counts <- table(open_mpg$model)


most_variations_model <- as.character(names(model_counts)[which.max(model_counts)])


print(paste("The manufacturer with the most models is:", most_models_manufacturer))
print(paste("The model with the most variations is:", most_variations_model))
```


2a.

```{r}
dfmpg <- data.frame(Manufacturer = open_mpg$manufacturer, Model = open_mpg$model)
unique_mpg <- unique(dfmpg)
unique_mpg

```

2b.

```{r}

library(ggplot2)

factoredManufacturer <- as.factor(unique_mpg$Manufacturer)
manufacturerPlot <- plot(as.factor(factoredManufacturer),
                 main = "Unique Model per Manufacturer",
                 cex.names = 0.8, las = 2)



library(dplyr)

unique_count <- unique_mpg %>%
  count(unique_mpg$Manufacturer)
unique_count

names(unique_count) <- c("Manufacturer", "Number of Models")
unique_count

ggplot(unique_count, aes(x = Manufacturer, y = `Number of Models`)) +
geom_bar(stat = "identity", fill = "violet")
```


2.Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
2a.What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
library(ggplot2)
ggplot(open_mpg, aes(model, manufacturer)) + geom_point()
```
ans. Plots the mpgdataset with model on the x-axis and manufacturer on the y-axis. Each point represents a car model

2b.
Knowing the number of models produced by each company is helpful.
I can add a label using the cty variable to each point and color code each point based on the class variable to make it more informative.


3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.

```{r}
library(ggplot2)

Aobj <-head(open_mpg, 20)
Aobj

ggplot(data = Aobj, aes(x = year, y = model, color = as.factor(model))) +
 geom_point(alpha = 0.7) +
 labs(title = "Relationship between Model, Year, and Manufacturer",
       x = "Year",
       y = "Model",
       color = "Manufacturer") +
 theme_minimal()

```

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.

```{r}
library(dplyr)
library(ggplot2)

head(open_mpg, 20)


open_mpg %>%
 group_by(model) %>%
 summarize(count = n()) -> grouped_open_mpg


print(grouped_open_mpg)

```


4a. Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels and colors. Show code and results.

```{r}

top_20 <- open_mpg %>%
 group_by(model) %>%
 summarise(count = n()) %>%
 arrange(desc(count)) %>%
 head(20)


plot <- ggplot(top_20, aes(x = reorder(model, count), y = count, fill = model)) +
 geom_bar(stat = "identity", color = "black") +
 scale_fill_brewer(palette = "Set2") +
 coord_flip() +
 labs(title = "Top 20 Most Popular Car Models",
       x = "Model",
       y = "Number of Cars") +
 theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))


print(plot)

```

4b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

```{r}
open_mpg <- data.frame(model = sample(letters[1:20], 1000, replace = TRUE))


top_20 <- open_mpg %>%
 group_by(model) %>%
 summarise(count = n()) %>%
 arrange(desc(count)) %>%
 head(20)

top_20 <- top_20 %>%
 group_by(model) %>%
 head(8)

plot <- ggplot(top_20, aes(x = reorder(model, count), y = count, fill = model)) +
 geom_bar(stat = "identity", color = "black") +
 scale_fill_brewer(palette = "Set2") +
 coord_flip() +
 labs(title = "Top 20 Most Popular Car Models",
       x = "Model",
       y = "Number of Cars") +
 theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

print(plot)

```

5.Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

```{r}
open_mpg <- data.frame(cyl = sample(6:10, 1000, replace = TRUE),
                 displ = sample(80:400, 1000, replace = TRUE),
                 engine_displacement = sample(200:4000, 1000, replace = TRUE))


plot <- ggplot(open_mpg, aes(x = cyl, y = displ, color = engine_displacement)) +
 geom_point() +
 scale_color_gradient(low = "blue", high = "red") +
 labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "No. of Cylinders",
       y = "Engine Displacement") +
 theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5))

print(plot)
```
5a.
To describe the relationship, we can analyze the scatter plot.

The plot shows that there is a downward trend. As the number of cylinders increases, the engine displacement tends to decrease.
There are also outliers, where the engine displacement is significantly larger or smaller than the others for the same number of cylinders.

```{r}
plot <- ggplot(open_mpg, aes(x = displ, y = cyl, color = "red")) +
 geom_point() +
 labs(title = "Relationship between Engine Displacement and Highway Mpg",
      x = "Engine Displacement",
      y = "Highway Mpg",
      color = "Highway Mpg") +
 theme_minimal()

print(plot)
```
